---
title: 'Demonstration TRS with R'
author: ""
date: "14 1 2021"
output:
  powerpoint_presentation:
    reference_doc: template.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install packages and test data

- Install package if not already installed

```{r,eval=FALSE}
# install recordSwapping from github
devtools::install_github("sdcTools/recordSwapping", build_vignettes = TRUE)

# install data.table 
# ~ generally good data wrangling package
install.packages(data.table)
```


## Load packages

- Do this for every new session in R ~ if you start working with R


```{r}
# load package
library(data.table)
library(recordSwapping)
```

## News, documentation and help page

- Current and development versions are on

    https://github.com/sdcTools/recordSwapping

- Documentation inside R


```{r,eval=FALSE}
# more detailed explanations/documentation:
vignette("recordSwapping") 

# documentation of core function
?recordSwap() 
```


## Load test data

```{r}
dat <- fread("test_data_10k.csv.gz")
class(dat) # class data.table
```

- View data in terminal 

```{r, eval=FALSE}
print(dat) # not printed to save space on slide
```

- Use data viewer

```{r, eval=FALSE}
View(dat)
```


## Apply targeted record swapping

1. Data preparation 

2. Set parameter for swapping

3. Apply record swapping

4. Compare output with original data 


## Data preparation

- Data needs to contain only integer/numeric columns 

    - no decimal places

- Columns which are not integer

```{r}
head(dat[,.(AGE.M,L001000)],3)
```

## Data preparation (2)

```{r}
str(dat[,.(AGE.M,L001000)])
```


- Solve programmatically 

```{r}
colType <- sapply(dat,class)
# filter colnames to non-integer columns
colnames(dat)[colType!="integer"]
```


## Data preparation (3)

- Convert these column to integer

```{r}
# extract X and Y coordinates
dat[,Y_coord:=as.integer(substr(L001000,5,8))]
dat[,X_coord:=as.integer(substr(L001000,10,13))]
dat[1,.(L001000,X_coord,Y_coord)]
```


```{r}
dat[,L001000:=NULL] # drop column
```

- Convert column to integer using `factor`

```{r}
dat[,AGE.M_old:=AGE.M]
age_levels <- levels(cut(1:101,breaks=c(seq(0,100,5),Inf),include.lowest = TRUE,right=FALSE))
dat[,AGE.M:=as.integer(factor(AGE.M,levels=age_levels))]
dat[,.N,by=.(AGE.M_old,AGE.M)][order(AGE.M)]
```

## Data preparation (4)

- Truncate household size

```{r}
dat[,Size:=pmin(5,Size)]
dat[!duplicated(HID),.N,by=.(Size)][order(Size)]
```


## Set parameter

- Necessary parameter (apart from data itself)

```{r}
# geographic hierarchy 
# read from left to right
# NUTS1 > NUTS2 > ...
hierarchy <- c("NUTS1","NUTS2")

# hid column
hid <- "HID"
```

## Set parameter (2)

- Variables for internal risk calculation
    
      - custom risk will be included in future releases

```{r}
# risk variables
risk_variables <- c("COC.M","POB.M")
k_anonymity <- 3
```


## Set paramter (3)

- Swaprate, similarity profile and seed for RNG

```{r}
swaprate <- 0.05
similar <- "Size"
seed <- 202103
```


## Apply record swapping

```{r}
dat_swapped <- recordSwap(data = dat, hid = hid,
                          hierarchy = hierarchy,
                          similar = similar,
                          risk_variables = risk_variables,
                          k_anonymity = k_anonymity,
                          swaprate = swaprate,
                          return_swapped_id = TRUE,
                          seed = seed)
```

- `return_swapped_id = TRUE` \rightarrow get household ID of swapped household



## Look at output

```{r, eval=FALSE}
View(dat_swapped)
print(dat_swapped)
```

- Number of swapped households

```{r}
dat_swapped[HID!=HID_swapped,uniqueN(HID)]
```


## Look at output (2)

- Table by NUTS1 and COC.L

```{r}
# original data
dat[,table(NUTS1,COC.L)]
```

```{r}
# swapped data
dat_swapped[,table(NUTS1,COC.L)]
```


## Information loss (1)

- Calculate counts of table cells

```{r}
tab_vars <- c("NUTS2","SEX","AGE.M","COC.H")
tab1 <- dat[,.(N1=.N),by=c(tab_vars)]
tab2 <- dat_swapped[,.(N2=.N),by=c(tab_vars)]

tab <- merge(tab1,tab2, all=TRUE)
tab[is.na(N1),N1:=0]
tab[is.na(N2),N2:=0]
```



## Information loss (2)

- Estimate information loss for each NUTS2 region

```{r}
# mean absolute deviation
ad <- tab[,mean(abs(N1-N2)),by=.(NUTS2)]
ad
```


## Information loss (3)

- Take average over summary statistics of each region

```{r}
ad[,mean(V1)]
```


- Calculate statistics on relative absolute difference (RAD)

    - use subtable for NUTS2, SEX and AGE.M due to small cell count

```{r}
# aggregate NUTS2, SEX and AGE.M
rad <- tab[,.(N1=sum(N1),N2=sum(N2)),by=.(NUTS2,SEX,AGE.M)]
```

## Information loss (4)

```{r}
# RAD by NUTS2
rad <- rad[,sum(abs(N1-N2)/N1),by=.(NUTS2)]
# take average
rad[,mean(V1)]
```


- Apply same logic to hellingers distance

```{r}
# hellingers distance
hd <- tab[,sqrt(1/2*sum((sqrt(N1)-sqrt(N2))^2)),
                 by=.(NUTS2)]
hd[,mean(V1)]
```


- see also R package `cellKey`, `?cellKey::ck_cnt_measures()` 




```{r, include=FALSE}
## ## Risk estimation

## - Proportion of unperturbed cells with cell count < 3  

## tab[,sum(N1<3&N2==N1)/sum(N1<3)]
```


## Multiple similarity variables

- Supply multiple variables to similarity profile to add conditions for swapping

```{r}
similar <- c("Size","HST")
```

```{r}
dat_swapped2 <- recordSwap(data = dat, hid = hid,
                          hierarchy = hierarchy,
                          similar = similar,
                          risk_variables = risk_variables,
                          k_anonymity = k_anonymity,
                          swaprate = swaprate,
                          return_swapped_id = TRUE,
                          seed = seed)
```

## Multiple similarity variables (2)

- Check Variable `HST` for swapped households

```{r}
dat_tph <- unique(dat[,.(HID_swapped=HID,HST_swapped=HST)])
dat_swapped[dat_tph,HST_swapped:=HST_swapped,on=.(HID_swapped)]

select_vars <- c("NUTS1", "NUTS2", "HID", "HID_swapped",
                 "Size", "HST", "HST_swapped")

head(dat_swapped[!duplicated(HID)][HST!=HST_swapped,..select_vars],4)
```


## Multiple similarity variables (2)

- Supply multiple variables to similarity profile to add conditions for swapping

```{r}
similar <- c("Size","HST")
```

```{r}
dat_swapped2 <- recordSwap(data = dat, hid = hid,
                          hierarchy = hierarchy,
                          similar = similar,
                          risk_variables = risk_variables,
                          k_anonymity = k_anonymity,
                          swaprate = swaprate,
                          return_swapped_id = TRUE,
                          seed = seed)
```


## Multiple similarity variables (3)

- Check again variable `HST` for swapped households

```{r}
dat_tph <- unique(dat[,.(HID_swapped=HID,HST_swapped=HST)])
dat_swapped2[dat_tph,HST_swapped:=HST_swapped,on=.(HID_swapped)]

select_vars <- c("NUTS1", "NUTS2", "HID", "HID_swapped",
                 "Size", "HST", "HST_swapped")


# HST identical for every swap
dat_swapped2[,all(HST==HST_swapped)]
```



```{r, include=FALSE}
similar <- c("Size")
```

## Carry along variables

```{r}
hids_swapped <- dat_swapped[HID!=HID_swapped]
head(hids_swapped[!duplicated(HID),.(NUTS1,NUTS2,NUTS3,LAU2)],4)
```

- Geographic variables are mixed up !?

## Carry along variables (2)

- Use parameter `carry_along` to carry geographic variables along while swapping

```{r}
carry_along <- c("NUTS3","LAU2","X_coord","Y_coord")
```


```{r}
dat_swapped3 <- recordSwap(data = dat, hid = hid,
                          hierarchy = hierarchy,
                          similar = similar,
                          risk_variables = risk_variables,
                          k_anonymity = k_anonymity,
                          swaprate = swaprate,
                          carry_along = carry_along,
                          return_swapped_id = TRUE,
                          seed = seed)
```


## Carry along variables (3)

- Swapping procedure is identical but geographic variables are coherent now

```{r}
hids_swapped3 <- dat_swapped3[HID!=HID_swapped]
head(hids_swapped3[!duplicated(HID),.(NUTS1,NUTS2,NUTS3,LAU2)])
```




